A cascaded supervised learning approach to inverse reinforcement learning

### 2013-03-29 piot2012cascading/ECML13_cascading.tex
Fin corrections Matthieu.

% Created 2012-10-15 lun. 14:49
\documentclass[smallextended]{svjour3}
\smartqed 
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{fixltx2e}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{soul}
\usepackage{textcomp}
\usepackage{marvosym}
\usepackage[integrals]{wasysym}
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{subfigure}
\tolerance=1000
%\usepackage{amsmath}
%\usepackage{amsthm}
%\newtheorem{definition}{Definition}
%\newtheorem{theorem}{Theorem}
\newtheorem{cor}{Corollary}
\usepackage{dsfont}
\usepackage{mathtools}
\mathtoolsset{showonlyrefs=true}
\providecommand{\alert}[1]{\textbf{#1}}
\newcommand{\E}{\mathbf{E}}
\newcommand{\argmax}{\operatorname*{argmax}} %\operatorname* pour les op. pouvant admettre des limites...
\newcommand{\Card}{\operatorname*{Card}} %\operatorname* pour les op. pouvant admettre des limites...
\begin{document}
\title{A cascaded supervised learning approach to inverse reinforcement learning}
\author{Edouard Klein$^{1,2}$ \and Bilal Piot $^{2,3}$\and Matthieu Geist $^{2}$\and Olivier Pietquin$^{2,3}$}
\titlerunning{Cascading approach to IRL}
\institute{
 1. ABC Team\\
 LORIA-CNRS, France.\\
\and
2. Supélec\\
 IMS-MaLIS Research group, France\\
 \texttt{firstname.lastname@supelec.fr}\\
\and 
3. UMI 2958 (GeorgiaTech-CNRS)\\
France\\
}
\date{\today}

@@

IRL is one of many ways to perform Apprenticeship Learning (AL): imitating a demonstrated expert policy, without necessarily explicitely looking for the reward function. The reward function nevertheless is of interest in its own right. As mentionned in \cite{russell1998learning}, its semantics can be analyzed in biology or econometrics for instance. Practically, the reward can be seen as a succinct description of a task. Discovering it removes the coupling that exists in AL between understanding the task and learning how to fulfill it. IRL allows the use of (A)DP or RL techniques to learn the task from the computed reward function. Although some approaches use it as a byproduct, most AL algorithm makes no use of the reward function, or have nothing to do with IRL altogether. A very straightforward non-IRL way to do AL is for example to use a multi-class classifier to directly learn the expert policy. We provide in the experiments (Sec.~\ref{sec:experiments}) a comparison between AL and IRL algorithms by using IRL as a way to do AL.

@@

%% Feel good comment
\section{Experiments}
\label{sec:experiments}
In this section, we empirically demonstrate the behavior of our approach. We begin by providing information pertaining to both benchmarks: an explanation about the amount and source of the available data, the rationale behind the heuristics we use to compensate for the dire data scarcity and a quick word about the contenders CSI is compared to are given Sec.~\ref{subsec:generalities}. We supply quantitative results and comparisons of CSI with state-of-the art approaches on first a classical RL benchmark (the mountain car) Sec.~\ref{subsec:mountaincar} and then on a more difficult highway driving simulator (Sec.~\ref{subsec:highway}).
\subsection{Generalities}
\label{subsec:generalities}
\subsubsection{Data scarcity}
\label{subsubsec:data}
The CSI algorithm was designed not to repeatedly solve the RL problem. This feature makes it particularly well-suited to environments where sampling the MDP is difficult or costly. In the experiments, CSI is fed only with data sampled according to the expert policy. This corresponds for example to a situation where a costly system can only be controlled by a trained operator as a bad control sequence could lead to a system breakdown.

@@

Our complete instanciation of CSI is summed up in Alg.~\ref{alg:CSIimp}.
\begin{algorithm}%[H]
    %\small
  \caption{A CSI instanciation with heuristics}
  \label{alg:CSIimp}
  \emph{\textbf{Given}} a dataset  $D_E = (s_k,a_k=\pi_E(a_k),s'_k)_k$\;\\
  \emph{\textbf{Construct}} the dataset $D_C=\{(s_i=s_k,a_i=\pi^E(s_i))=a_k\}$ \;\\
  \emph{\textbf{Train}} a score function-based classifier on $D_C$, obtaining decision rule $\pi_C$ and score function $q:S\times A \rightarrow \mathbb R$ \;\\
\emph{\textbf{Construct}} the dataset $\{((s_{j}=s_k,a_{j}=a_k),\hat r_j)_j\}$ with $\hat r_j = q(s_j,a_j) - \gamma q(s'_j=s'_k,\pi_C(s'_j=s'_k))$\;\\
\emph{\textbf{Set}} $r_{min} = \min_j\hat r_j - 1.$\;\\
\emph{\textbf{Construct}} the training set  $D_R = \{((s_{j}=s_k,a_{j}=a_k),\hat r_j)_j\}\cup\{((s_j=s_k,a),r_{min})_{j;\forall a\neq \pi_E(s_j) = a_k}\}$ \;\\
\emph{\textbf{Learn}} a reward function $\hat R^C$ from the trainsing set $D_R$\;\\
\emph{\textbf{Output}} the reward function $\hat R^{C}: (s,a) \mapsto \omega^T \phi(s,a)$ \;
\end{algorithm}
\subsection{Mountain car}
\label{subsec:mountaincar}
The mountain car is a classical toy problem in RL: an underpowered car is tasked with climbing a steep hill. In order to do so, it has to first move away from the target and climb the slope on its left, and then it moves right, gaining enough momentum to climb the hill on the right on top of which lies the target. We used standard parameters for this problem, as can for example be found in \cite{sutton1998reinforcement}. When training an RL agent, the reward is, for example, $1$ if the car's position is greater than $0.5$ and $0$ anywhere else. The expert was a very simple hand crafted policy in which the agent uses the power of the car to go in the direction it already moves ({\it i.e.} go left when the speed is negative, right when it is positive)\\

The initial position of the car was uniformly randomly picked in $[-1.2;-0.9]$ and its speed uniformly randomly picked in $[-0.07;0]$. From this position, the hand-crafted policy was left to play until the car reached the objective ({\it i.e.} a position greater than $0.5$) at which point the episode ended. Enough episodes were played (and the last one was truncated) so that the dataset $D_E$ contained exactly $n$ samples, with $n$ successively equal to $10$, $30$, $100$ and $300$. With these parameters, the expert is always able to reach the top on the first time it tries to climb the hill on the right. Therefore a whole part of the state space (when the position is on the hill on the right and the speed is negative) is not visited by the expert. This hole about the state space in the data will be dealt with differently by the classifier and the IRL algorithms. The classifier will use its generalization power to find a default action in this part of the state space, while the IRL algorithms will devise a default reward; a (potentially untrained) RL agent finding itself in this part of the state space will use the reward signal to decide what to do, making use of new data available at that time.

@@

The objective criterion for success is the number of steps needed to attain the goal when starting from a state picked at random ; the lesser the better. We can see Fig.~\ref{fig:MountainCar} that the optimal policies for the rewards found by SCIRL and CSI very rapidly attain expert-level performance and outperform the optimal policy for the reward of RE and the classification policy.
\begin{figure}
  \includegraphics[width=\textwidth]{Exp11.pdf}
  \caption{Performance of various policies on the mountain car problem. This is the mean over 50 runs.}
  \label{fig:MountainCar}
\end{figure}

The classifier for CSI was an off-the-shelf SVM\footnote{\url{http://scikit-learn.org/stable/modules/generated/sklearn.svm.SVC.html}}, the regressor was an off-the-shelf SVR\footnote{\url{http://scikit-learn.org/stable/modules/generated/sklearn.svm.SVR.html}}.

@@

Results are shown Fig.~\ref{fig:Highway} and Tab.~\ref{tab:highway}. We give the values of $\mathbf{E}_{s\sim\mathcal{U}}[V^{\pi}_{R_E}(s)]$ with $\pi$ being in turn the optimal policy for the rewards given by SCIRL, CSI and RE, the policy $\pi_C$ of the classifier (the very one the classification step of CSI outputs), and the optimal policy for a randomly drawn reward. Performance for CSI is slightly but definitively higher than for SCIRL (see the variance in Tab.~\ref{tab:highway}), slightly below the performance of the expert itself. Very few samples (100) are needed to reliably achieve expert-level performance.

@@

Furthermore, the classifier whose results are diplayed here is the output of the first step of the algorithm. The classification performance is obviously not that good, this hints to the fact that our algorithm may be empirically more forgiving of classification errors than our theoretical bound lets us expect.
\begin{figure}
      \begin{tabular}{|l|l|l|l|l|}
        \hline
        &\bf CSI& \bf SCIRL& \bf RE& \bf Classif\\
        \hline \multicolumn{5}{|c|}{$n = 9$}\\ \hline
        Var& 5.139& 5.684& 4.686& 0.045 \\
        Min& -2.708& -3.175& -2.717& -1.564 \\
        Mean& 2.419& 2.762& 2.286& -1.149 \\
        Max& 7.050& 7.060& 7.027& -0.596 \\
        \hline \multicolumn{5}{|c|}{$n = 49$}\\ \hline
        Var& 0.214& 0.517& 0.701& 0.121 \\
        Min& 5.406& 4.652& 4.214& -1.148 \\
        Mean& 7.012& 6.784& 6.314& -0.158 \\
        Max& 7.454& 7.529& 7.465& 0.619 \\
        \hline \multicolumn{5}{|c|}{$n = 100$}\\ \hline
        Var& 0.011& 0.269& 0.486& 0.154 \\
        Min& 7.078& 5.724& 5.487& -0.574 \\
        Mean& 7.364& 7.194& 6.778& 0.456 \\
        Max& 7.526& 7.562& 7.509& 1.483 \\ \hline
        \end{tabular}
      \begin{tabular}{|l|l|l|l|l|}
        \hline
        &\bf CSI& \bf SCIRL& \bf RE& \bf Classif\\
        \hline \multicolumn{5}{|c|}{$n = 225$}\\ \hline
        Var& 0.006& 0.029& 0.146& 0.167 \\
        Min& 7.172& 5.882& 5.612& 0.128 \\
        Mean& 7.459& 7.377& 7.215& 1.151 \\
        Max& 7.563& 7.572& 7.578& 2.220 \\
        \hline \multicolumn{5}{|c|}{$n = 400$}\\ \hline
        Var& 0.002& 0.002& 0.005& 0.267 \\
        Min& 7.321& 7.247& 7.144& 0.041 \\
        Mean& 7.495& 7.369& 7.279& 1.744 \\
        Max& 7.552& 7.557& 7.519& 2.761 \\
        \hline
      \end{tabular}
      \caption{Raw results on the Highway driving problem, see how the variance of SCIRL and CSI decreases.}
     \label{tab:highway}
      \end{figure}
  \begin{figure}
   \subfigure[Mean performance over 100 runs on the Highway driving problem.]{\label{tutu}\includegraphics[width=.5\linewidth]{"Exp14"}}
   \subfigure[Zoom of Fig~\ref{tutu} showing the ranking of the three IRL algorithms.]{\includegraphics[width=.5\linewidth]{"Exp14_zoom"}}
  \caption{Results on the highway driving problem.}
  \label{fig:Highway}

@@

\section{Conclusion}
\label{sec:conclusion}
We have introduced a new way to perform IRL by cascading two supervised approaches. The expert is theoretically shown to be near-optimal for the reward function the proposed algorithm outputs, given small classification and regression errors. Practical examples of classifiers and regressors have been given, and two combinations have been empirically (on two classic benchmarks) shown to be very resilient to dire lack of data on the input (only data from the expert was used to retrieve the reward function), provided the use of simple heuristics. On one of the benchmarks (a highway driving simulator), our algorithm is shown to outperform other state-of-the-art approaches. We plan on deepening the analysis of the theoretical properties of our approach and on applying it to real world robotics problems.

>>>>

% Created 2012-10-15 lun. 14:49
\documentclass[smallextended]{svjour3}
\smartqed 
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{fixltx2e}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{soul}
\usepackage{textcomp}
\usepackage{marvosym}
\usepackage[integrals]{wasysym}
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{subfigure}
% scientific notation, 1\e{9} will print as 1x10^9
\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}}
\usepackage{booktabs} % Fancy tables
\tolerance=1000
%\usepackage{amsmath}
%\usepackage{amsthm}
%\newtheorem{definition}{Definition}
%\newtheorem{theorem}{Theorem}
\newtheorem{cor}{Corollary}
\usepackage{dsfont}
\usepackage{mathtools}
\mathtoolsset{showonlyrefs=true}
\providecommand{\alert}[1]{\textbf{#1}}
\newcommand{\E}{\mathbf{E}}
\newcommand{\argmax}{\operatorname*{argmax}} %\operatorname* pour les op. pouvant admettre des limites...
\newcommand{\Card}{\operatorname*{Card}} %\operatorname* pour les op. pouvant admettre des limites...
\begin{document}
\title{A cascaded supervised learning approach to inverse reinforcement learning}
\author{Edouard Klein$^{1,2}$ \and Bilal Piot $^{2,3}$\and Matthieu Geist $^{2}$\and Olivier Pietquin$^{2,3}$}
\titlerunning{Cascading approach to IRL}
\institute{
 1. ABC Team\\
 LORIA-CNRS, France.\\
\and
2. Supélec\\
 IMS-MaLIS Research group, France\\
 \texttt{firstname.lastname@supelec.fr}\\
\and 
3. UMI 2958 (GeorgiaTech-CNRS)\\
France\\
}
\date{\today}

@@

IRL is one of many ways to perform Apprenticeship Learning (AL): imitating a demonstrated expert policy, without necessarily explicitely looking for the reward function. The reward function nevertheless is of interest in its own right. As mentionned in \cite{russell1998learning}, its semantics can be analyzed in biology or econometrics for instance. Practically, the reward can be seen as a succinct description of a task. Discovering it removes the coupling that exists in AL between understanding the task and learning how to fulfill it. IRL allows the use of (A)DP or RL techniques to learn the task from the computed reward function. A very straightforward non-IRL way to do AL is for example to use a multi-class classifier to directly learn the expert policy. We provide in the experiments (Sec.~\ref{sec:experiments}) a comparison between AL and IRL algorithms by using IRL as a way to do AL.

@@

%% Feel good comment
\section{Experiments}
\label{sec:experiments}
In this section, we empirically demonstrate the behavior of our approach. We begin by providing information pertaining to both benchmarks: an explanation about the amount and source of the available data, the rationale behind the heuristics we use to compensate for the dire data scarcity and a quick word about the contenders CSI is compared to are given Sec.~\ref{subsec:generalities}. We supply quantitative results and comparisons of CSI with state-of-the art approaches on first a classical RL benchmark (the mountain car) Sec.~\ref{subsec:mountaincar} and then on a highway driving simulator (Sec.~\ref{subsec:highway}).
\subsection{Generalities}
\label{subsec:generalities}
\subsubsection{Data scarcity}
\label{subsubsec:data}
The CSI algorithm was designed not to repeatedly solve the RL problem. This feature makes it particularly well-suited to environments where sampling the MDP is difficult or costly. In the experiments, CSI is fed only with data sampled according to the expert policy. This corresponds for example to a situation where a costly system can only be controlled by a trained operator as a bad control sequence could lead to a system breakdown.

@@

Our complete instanciation of CSI is summed up in Alg.~\ref{alg:CSIimp}.
\begin{algorithm}%[H]
    %\small
  \caption{A CSI instanciation with heuristics}
  \label{alg:CSIimp}
  \emph{\textbf{Given}} a dataset  $D_E = (s_k,a_k=\pi_E(a_k),s'_k)_k$\;\\
  \emph{\textbf{Construct}} the dataset $D_C=\{(s_i=s_k,a_i=\pi^E(s_i))=a_k\}$ \;\\
  \emph{\textbf{Train}} a score function-based classifier on $D_C$, obtaining decision rule $\pi_C$ and score function $q:S\times A \rightarrow \mathbb R$ \;\\
\emph{\textbf{Construct}} the dataset $\{((s_{j}=s_k,a_{j}=a_k),\hat r_j)_j\}$ with $\hat r_j = q(s_j,a_j) - \gamma q(s'_j=s'_k,\pi_C(s'_j=s'_k))$\;\\
\emph{\textbf{Set}} $r_{min} = \min_j\hat r_j - 1.$\;\\
\emph{\textbf{Construct}} the training set  $D_R = \{((s_{j}=s_k,a_{j}=a_k),\hat r_j)_j\}\cup\{((s_j=s_k,a),r_{min})_{j;\forall a\neq \pi_E(s_j) = a_k}\}$ \;\\
\emph{\textbf{Learn}} a reward function $\hat R^C$ from the trainsing set $D_R$\;\\
\emph{\textbf{Output}} the reward function $\hat R^{C}: (s,a) \mapsto \omega^T \phi(s,a)$ \;
\end{algorithm}
\subsection{Mountain car}
\label{subsec:mountaincar}
The mountain car is a classical toy problem in RL: an underpowered car is tasked with climbing a steep hill. In order to do so, it has to first move away from the target and climb the slope on its left, and then it moves right, gaining enough momentum to climb the hill on the right on top of which lies the target. We used standard parameters for this problem, as can for example be found in \cite{sutton1998reinforcement}. When training an RL agent, the reward is, for example, $1$ if the car's position is greater than $0.5$ and $0$ anywhere else. The expert was a very simple hand crafted policy in which the agent uses the power of the car to go in the direction it already moves ({\it i.e.}, go left when the speed is negative, right when it is positive).\\

The initial position of the car was uniformly randomly picked in $[-1.2;-0.9]$ and its speed uniformly randomly picked in $[-0.07;0]$. From this position, the hand-crafted policy was left to play until the car reached the objective ({\it i.e.}, a position greater than $0.5$) at which point the episode ended. Enough episodes were played (and the last one was truncated) so that the dataset $D_E$ contained exactly $n$ samples, with $n$ successively equal to $10$, $30$, $100$ and $300$. With these parameters, the expert is always able to reach the top on the first time it tries to climb the hill on the right. Therefore, a whole part of the state space (when the position is on the hill on the right and the speed is negative) is not visited by the expert. This hole about the state space in the data will be dealt with differently by the classifier and the IRL algorithms. The classifier will use its generalization power to find a default action in this part of the state space, while the IRL algorithms will devise a default reward; a (potentially untrained) RL agent finding itself in this part of the state space will use the reward signal to decide what to do, making use of new data available at that time.

@@

The classifier for CSI was an off-the-shelf SVM\footnote{\url{http://scikit-learn.org/stable/modules/generated/sklearn.svm.SVC.html}} which also was the classifier we evaluate, the regressor of CSI was an off-the-shelf SVR\footnote{\url{http://scikit-learn.org/stable/modules/generated/sklearn.svm.SVR.html}}. RE and SCIRL need features over the state space, we used the same evenly-spaced hand-tuned $7\times 7$ RBF network for both algorithms.

The objective criterion for success is the number of steps needed to attain the goal when starting from a state picked at random ; the lesser the better. We can see Fig.~\ref{fig:MountainCar} that the optimal policies for the rewards found by SCIRL and CSI very rapidly attain expert-level performance and outperform the optimal policy for the reward of RE and the classification policy. When very few samples are available, CSI does better than SCIRL (with such a low $p$-value for $n=10$, see Tab.~\ref{tab:mc}, the hypothesis that the mean performance is equal can be rejected); SCIRL catches up when more samples are available. Furthermore, CSI required very little engineering as we cascaded two off-the-shelf implementations whereas SCIRL used hand-tuned features and a custom classifier.
\begin{table}
	\begin{center}
		\begin{tabular}{cc}
		\toprule
			Number of expert samples & $p$-value\\
			\midrule
			$10$ & $4.5195\e{-14}$\\
			$30$ & $0.0131123$\\
			$100$ & $0.139975$\\
			$300$ & $0.893752$\\
		\bottomrule
		\end{tabular}
	\end{center}
        \caption{Kolmogorov–Smirnov mean equality test $p$-values for CSI and SCIRL on the mountain car. High values (>0.01) means that the hypothesis that the means are equal cannot be rejected.}
        \label{tab:mc}
\end{table}
\begin{figure}
  \includegraphics[width=\textwidth]{Exp11.pdf}
  \caption{Performance of various policies on the mountain car problem. This is the mean over 100 runs.}
  \label{fig:MountainCar}
\end{figure}

@@

Results are shown Fig.~\ref{fig:Highway}. We give the values of $\mathbf{E}_{s\sim\mathcal{U}}[V^{\pi}_{R_E}(s)]$ with $\pi$ being in turn the optimal policy for the rewards given by SCIRL, CSI and RE, the policy $\pi_C$ of the classifier (the very one the classification step of CSI outputs), and the optimal policy for a randomly drawn reward. Performance for CSI is slightly but definitively higher than for SCIRL (see the $p$-values for the mean equality test in Tab.~\ref{tab:highway}), slightly below the performance of the expert itself. Very few samples (100) are needed to reliably achieve expert-level performance.

@@

Furthermore, the classifier whose results are diplayed here is the output of the first step of the algorithm. The classification performance is obviously not that good, this hints to the fact that our algorithm may be empirically more forgiving of classification errors than our theoretical bound lets us expect.
\begin{table}[ht]
	\begin{center}
		\begin{tabular}{cc}
		\toprule
			Number of expert samples & $p$-value\\
			\midrule
			$9$ & $0.193042$\\
			$49$ & $0.00178467$\\
			$100$ & $0.00178467$\\
			$225$ & $6.66654\e{-22}$\\
			$400$ & $3.70654\e{-36}$\\
		\bottomrule
		\end{tabular}
	\end{center}
        \caption{Kolmogorov–Smirnov mean equality test $p$-values for CSI and SCIRL on the highway simulator. High values (>0.01) means that the hypothesis that the means are equal cannot be rejected.}
        \label{tab:highway}
\end{table}
  \begin{figure}
   \subfigure[Mean performance over 100 runs on the Highway driving problem.]{\label{tutu}\includegraphics[width=.5\linewidth]{"Exp14"}}
   \subfigure[Zoom of Fig~\ref{tutu} showing the ranking of the three IRL algorithms.]{\includegraphics[width=.5\linewidth]{"Exp14_zoom"}}
  \caption{Results on the highway driving problem.}
  \label{fig:Highway}

@@

\section{Conclusion}
\label{sec:conclusion}
We have introduced a new way to perform IRL by cascading two supervised approaches. The expert is theoretically shown to be near-optimal for the reward function the proposed algorithm outputs, given small classification and regression errors. Practical examples of classifiers and regressors have been given, and two combinations have been empirically (on two classic benchmarks) shown to be very resilient to dire lack of data on the input (only data from the expert was used to retrieve the reward function), provided the use of simple heuristics. On both benchmarks, our algorithm is shown to outperform other state-of-the-art approaches (although SCIRL catches up on the mountain car). We plan on deepening the analysis of the theoretical properties of our approach and on applying it to real world robotics problems.
